---
title: ""
author: ""
output: html_document
---

## Data exploration {.tabset}


Operations to Learn:

`select`: select specific columns of a data frame

`filter`: extract a subset of rows from a data frame based on logical conditions

`arrange`: reorder rows of a data frame

`rename`: rename variables in a data frame

`mutate`: add new variables/columns or transform existing variables

`group_by` : groups data by column factors

`summarise`: generate summary statistics of different variables in the data frame, possibly within strata


### Manipulation

We'll use the `dplyr` package to manipulate our data. `dplyr` and other packages are part of the `tidyverse` package.

```{r data tidying}
# Load the required package
library(tidyverse)

# Look at datasets available in R for practise
data()

df <- datasets::mtcars
?mtcars

## Select


# Namespaces


# now let's get data about the miles and transmission of the cars
dplyr::select(df, c(mpg,am))

?dplyr::select() # for help

# what if we want to select more than one column:
# Let's get data about the miles and transmission of the cars
dplyr::select(df, mpg:am)   # approach 1

dplyr::select(df, c(mpg, am, carb, disp)) # approach 3

dplyr::select(df, contains("c"))


## Filter
# we want cars with more than 20 miles per gallon
dplyr::filter(df, mpg > 20)


# filtering by multiple commands

dplyr::filter(df, vs == 0 & mpg > 20)  # for 'AND'


dplyr::filter(df, vs != 0 & mpg > 20) # not 'vs == 0' and mpg > 20


dplyr::filter(df, vs != 0 & (mpg > 20 | hp < 91)) # not 'vs == 0' and mpg > 20 OR hp < 91



## Arrange
# we want to order cars by weight in descending order
dplyr::arrange(df, desc(wt))



## Rename
# we want rename the engine column name
dplyr::rename(df, engine = vs)


## mutate
# we want to insert a new column showing the weight in kg instead of pounds
dplyr::mutate(df, wt_kg = wt * 0.45, .after = wt)


## group_by
dplyr::summarise(df, mean(mpg))


```


Piping: the “pipe” is used to connect multiple operations together into a pipeline
` %>% ` or `|>`

`Windows:` Ctrl + Shift + M    `Mac:` Cmd + Shift + M

```{r piping}
# Usage

df %>% dplyr::filter(...) %>% dplyr::select(...) %>% 
  dplyr::mutate() %>% dplyr::group_by() %>% dplyr::summarise()


```


```{r data manipulation exercise}
# Exercise: 
# Jump to powerpoint


```







### Wrangling/Reshaping

`separate ` `pivot_wider` `pivot_longer` `tibble` 

```{r wrangling}

####---Separate
# separating a column into mmultiple colums
df <- read.csv("flow_data.csv")  # read in data


####---- Pivot longer
df %>% pivot_longer(col=c(2:4), names_to = "Subject", values_to = c("Scores")) # OR
df %>% pivot_longer(col=c(Math:Arts), names_to = "Subject", values_to = c("Scores"))


####---- Pivot_wider
df %>% pivot_wider(names_from = "Subject", values_from = "Scores")






####---- Joining tables






####---- Moving columns to rownames and vice versa




```



