---
title: ""
author: ""
output: html_document
---


## Basic R Syntax {.tabset}

### Variable Assignment and Basic Operations

\<- = assignment operator. Shortcut in `Windows/Linux` is 'Alt + -". For `Mac`, use Option + -\`

```{r variables and operations, message=FALSE, warning=FALSE}
# Variable assignment
a <- 20
b <- -17
c <- 21 

# Basic Operations
x <- a + b

y <- x * c

z <- y - a^2


# remove objects from an the environment
rm(a,b,c,x,y,z)

```

### Data types

Let's take a look at some R data types `integers, numneric, logical, character data types`

```{r Data types, message=FALSE, warning=FALSE}
# Integer data type represents whole numbers without decimal places.
a <- 20L
b <- -3L
c <- 7 # without the 'L', R stores it as an integer/double (7.00)


# Numeric data type represents decimal or floating-point numbers
x <- 10.5
y <- -3.2

# Logical data type represents Boolean values (TRUE or FALSE).
is_true <- TRUE
is_false <- FALSE

# Character data type represents strings of characters.
name <- "Princess"
age <- "Twenty"
age2 <- "20"

# Want to know what data type you have?
# Using typeof()
typeof(a)
typeof(age2)
typeof(c)
typeof(is_true)

# Using str()
str(a)
str(age2)

```

### Data structures

Let's take a look at data structures in R `vectors, matrices, lists, and data frames`

```{r Vectors}
## A vector is a one-dimensional array that can hold elements of the same data type.
a <- c("apple", "banana", "orange")
b <- c(1, 2, 3, 4, 5)
c <- c(-2, 10, 3)

## naming a vector
names(b) <- c("a","b","c","d","e")

## Working with vectors: indexing
a[3]  # selects the third element

b[-4] # selects everything in 'b' except the fourth element

b[1:3] # selects elements from 1 to 3 in b

2 * b  # doubles every element in b

# What will the output of this? Go to mentimeter
7 * b[1] + b[2:4]

```

```{r matrix}
## A matrix is a two-dimensional data structure with rows and columns.
a <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 2, ncol = 3)
a[2,3]

## What's the outcome? Go to mentimeter
a[2,1] * a[1,3] 

## Combining matrices: 
# by colum = cbind()
# by rows = rbind()

b <- matrix(seq(10:20), nrow = 3, ncol = 3)

c <- rbind(a,b)


# Will rbind work for a and b?

```

```{r lists}
## A list is a collection of objects that can be of different data types.
a <- list(name = "Paul", age = 25, city = "Oslo")

a$name
a$city

# Adding something new??
c <- append(a, "new", after = 3) 
d <- append(a, list(new="new"), after = 3) # give a name to the thing

# Removing something
d$new <- NULL

```

```{r dataframes}
## A data frame is a type of list with tabular data structure.
a <- data.frame(name = c("John", "Jane", "Mike"),
                age = c(25, 30, 35),
                city = c("New York", "London", "Paris"))
View(a)

names(a)   # tells you column names of the dataframe
colnames(a) # does same
rownames(a) # tells you the rownames of the dataframe
a$name  # selects the column called 'name'
a[3] # indexes the third colum

a[3,] # what does this do?

a[1,3] # ?

```

```{r factors}
# Factors
# Factors are used to represent categorical data with fixed levels.
b <- factor(c("Male", "Female", "Male"), levels = c("Male", "Female"))

# Let's make city in dataframe 'a' a factor
a$city <- as.factor(a$city)
a$city

a$city <- factor(a$city, levels = c("Paris", "New York", "London"))
a$city

# Want to know what data structures you have?
# class() # what kind of object is it?
# typeof() # what is the object's data type?
# length() # how long is it?
# attributes() # does it have any metadata?

```


### Importing and exporting data

`read.table`, `read.csv`, `read.delim` for reading tabular data `load`, for reading in saved workspaces

```{r import}
# Set-up working directories
setwd("C:/Users/ebenezer.asiedu/Desktop/Grad_Sch/PhD/training/R-miniWorkshop")

df <- read.table("import.txt")

# Let's make the first rows the column names
df <- read.table("import.txt", header = T)

# Look for help on how to use read.table()
??read.table


df <- read.csv("import.csv", header = T)

```

`write.table`, `write.csv`, `write.delim` for writing tabular data `save`, for saving R workspaces

```{r export}
# Set-up working directories
setwd("C:/Users/ebenezer.asiedu/Desktop/Grad_Sch/PhD/training/R-miniWorkshop")

write.csv(df, "written.csv")

```
