---
title: ""
author: ""
output: html_document
---

## Data exploration {.tabset}

### Manipulations

We'll use the `dplyr` package to manipulate our data. `dplyr` and other packages are part of the `tidyverse` package.

Operations to Learn:

`select`: return a subset of the columns of a data frame, using a flexible notation

`filter`: extract a subset of rows from a data frame based on logical conditions

`arrange`: reorder rows of a data frame

`rename`: rename variables in a data frame

`mutate`: add new variables/columns or transform existing variables

`group_by` : groups data by column factors

`summarise`: generate summary statistics of different variables in the data frame, possibly within strata

`separate ` `pivot_wider` `pivot_longer`

```{r data tidying}
# Load the required package
library(tidyverse)

# Look at datasets available in R for practise
data()

df <- datasets::mtcars
?mtcars

## Select
# Let's get data about the miles and transmission of the cars
dplyr::select(df, mpg:am)

# now let's get data about the miles and transmission of the cars
dplyr::select(df, c(mpg,am))

## Filter
# we want cars with more than 20 miles per gallon
dplyr::filter(df, mpg > 20)

## Arrange
# we want to order cars by weight in descending order
dplyr::arrange(df, desc(wt))


## Rename
# we want rename the engine column name
dplyr::rename(df, engine = vs)


## mutate
# we want to insert a new column showing the weight in kg instead of pounds
dplyr::mutate(df, wt_kg = wt * 0.45, .after = wt)


## group_by
dplyr::summarise(df, mean(mpg))

```

### Piping

%>%: the “pipe” is used to connect multiple operations together into a pipeline

```{r}



```


### Descriptive Statistics

how to quikly look at basic statistics of your data

```{r descriptive stats}



```

